|

### 2023-07-17 paper/02-data.Rmd
fixed abbreviations (#14)

Table \ref{tab:track} displays a tracking data example for a [play](https://www.raiders.com/video/de-maxx-crosby-sacks-qb-teddy-bridgewater-for-a-loss-of-6-yards-nfl) from the 2021 NFL regular season week six matchup between the Las Vegas Raiders and Denver Broncos, which ends with Broncos QB Teddy Bridgewater getting sacked by Raiders DE Maxx Crosby[^crosbysack].
In addition, Figure \ref{fig:fig_field_raw} presents the locations of every Las Vegas (in black) and Denver (in orange) player on the field from this play at 1, 2, 3 and 4 seconds after the ball snap, with Maxx Crosby highlighted in blue.

@@

\begin{table}
\caption{Example of tracking data for a play during the Las Vegas Raiders versus Denver Broncos NFL game on October 17, 2021. The data shown here are for Raiders DE Maxx Crosby, and the frames included are between the ball snap and when the sack by Crosby is recorded. \label{tab:track}}
\centering
\begin{tabular}{ccccccccc}
\hline
frameId & x & y & s & a & dis & o & dir & event \\ 
\hline
7 & 67.68 & 29.89 & 0.34 & 1.57 & 0.04 & 124.86 & 88.21 & ball\_snap \\ 
8 & 67.76 & 29.89 & 0.69 & 2.13 & 0.08 & 124.07 & 89.59 & None \\ 
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\ 
50 & 73.67 & 25.06 & 4.19 & 2.62 & 0.42 & 134.21 & 125.26 & qb\_sack \\ 
\hline
\end{tabular}
\end{table}

```{r}
#| label: fig_field_raw
#| fig.cap: "A display of the player tracking data for a play during the Las Vegas Raiders (defense, in black) versus Denver Broncos (offense, in orange) NFL game on October 17, 2021. Raiders DE Maxx Crosby is highlighted in blue. Snapshots are captured at 1, 2, 3, and 4 seconds after the ball snap."
#| fig.width: 6.6
#| fig.height: 4.4
source(here("scripts", "11_fig_field_raw.R"))
fig_field_raw
```

>>>>

Table \ref{tab:track} displays a tracking data example for a [play](https://www.raiders.com/video/de-maxx-crosby-sacks-qb-teddy-bridgewater-for-a-loss-of-6-yards-nfl) from the 2021 NFL regular season week six matchup between the Las Vegas Raiders and Denver Broncos, which ends with Broncos quarterback Teddy Bridgewater getting sacked by Raiders defensive end Maxx Crosby.
In addition, Figure \ref{fig:fig_field_raw} presents the locations of every Las Vegas (in black) and Denver (in orange) player on the field from this play at 1, 2, 3 and 4 seconds after the ball snap, with Maxx Crosby highlighted in blue.

@@

\begin{table}
\caption{Example of tracking data for a play during the Las Vegas Raiders versus Denver Broncos NFL game on October 17, 2021. The data shown here are for Raiders defensive end Maxx Crosby, and the frames included are between the ball snap and when the sack by Crosby is recorded. \label{tab:track}}
\centering
\begin{tabular}{ccccccccc}
\hline
frameId & x & y & s & a & dis & o & dir & event \\ 
\hline
7 & 67.68 & 29.89 & 0.34 & 1.57 & 0.04 & 124.86 & 88.21 & ball\_snap \\ 
8 & 67.76 & 29.89 & 0.69 & 2.13 & 0.08 & 124.07 & 89.59 & None \\ 
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\ 
50 & 73.67 & 25.06 & 4.19 & 2.62 & 0.42 & 134.21 & 125.26 & qb\_sack \\ 
\hline
\end{tabular}
\end{table}

```{r}
#| label: fig_field_raw
#| fig.cap: "A display of the player tracking data for a play during the Las Vegas Raiders (defense, in black) versus Denver Broncos (offense, in orange) NFL game on October 17, 2021. Raiders defensive end Maxx Crosby is highlighted in blue. Snapshots are captured at 1, 2, 3, and 4 seconds after the ball snap."
#| fig.width: 6.6
#| fig.height: 4.4
source(here("scripts", "11_fig_field_raw.R"))
fig_field_raw
```

### 2023-07-17 paper/04-results.Rmd
fixed abbreviations (#14)

```{r}
#| label: fig_crosby_curves
#| fig.cap: "Changes in STRAIN, velocity, and distance from the QB for Maxx Crosby over the course of a successful pass rush play that results in a sack. The STRAIN for other pass rushers throughout this play is also displayed."
#| fig.width: 4.8
#| fig.height: 2.7
source(here("scripts", "02_fig_crosby_curves.R"))
fig_crosby_curves
```

@@

We observe that for the first two seconds, Crosby is being blocked by a Denver offensive lineman and unable to get close to the quarterback, hence the corresponding STRAIN values are virtually zero. 
Suddenly at around three seconds after the snap, the STRAIN for Crosby starts to increase after the Raiders DE is freed up and charges towards Bridgewater. 
At 4 seconds after the snap, Crosby's STRAIN is 2.30, which means at his current moving rate, it will take Crosby about 0.43 (1/2.30) seconds to make the distance between him and the quarterback 0 (i.e. essentially sack the quarterback).
This matches well with the final outcome of the play, as the sack takes place at the very last frame (4.4 seconds) where the estimated STRAIN for Crosby reaches its peak at 3.96.

@@

```{r}
#| label: fig_crosby_curves_unsucess
#| fig.cap: "Changes in STRAIN, velocity, and distance from the QB for Maxx Crosby over the course of an unsuccessful pass rush play. The STRAIN for other pass rushers throughout this play is also displayed."
#| fig.width: 4.8
#| fig.height: 2.7
source(here("scripts", "09_fig_play_unsuccess.R"))
fig_crosby_curves_unsucess
```

<!-- ```{r} -->
<!-- #| label: fig_features_combined -->
<!-- #| fig.cap: "Changes in STRAIN, velocity, and distance from the QB for Maxx Crosby over the course of a successful pass rush play that results in a sack (left) and an unsuccessful pass rush play which results in a long pass attempt (right). The STRAIN for other pass rushers throughout each play is also displayed." -->
<!-- #| fig.width: 6 -->
<!-- #| fig.height: 3.6 -->
<!-- source(here("scripts", "15_fig_features_combined.R")) -->
<!-- fig_features_combined_bw -->
<!-- ``` -->

@@

```{r}
#| label: fig_pos_curves
#| fig.cap: "STRAIN curves for different positions. Edge rushers (OLB and DE) tend to generate more STRAIN than interior rushers (DT and NT). The dotted gray line represents the average STRAIN curve for all players without accounting for position."
#| fig.width: 4.5
#| fig.height: 3
source(here("scripts", "03_fig_pos_curves.R"))
fig_pos_curves
```

@@

Figure \ref{fig:fig_pos_curves} displays the average STRAIN by position for the first 40 frames (4 seconds) after the snap.
For each position, the curve is based on the average of the observed STRAIN at each frame across all plays and players within the position.
We observe a clear difference in STRAIN between edge rushers (OLBs and DEs) and interior linemen (DTs and NTs).
Specifically, edge rushers have higher STRAIN than interior linemen on average, as they are more easily able to approach the quarterback on the edge of the pocket versus, for instance, a nose tackle attacking the line head on.

@@

Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size, as discussed in Section \ref{sec:defn}.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass rushers (NTs and DTs) separately from edge rushers (OLBs and DEs).
Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season.
The tables also consist of the total number of hits, hurries, and sacks (determined from PFF scouting data) for each defender.

@@

<!-- From @Franks2016Meta (github https://github.com/afranks86/meta-analytics) -->

<!--

Discrimination: Does the metric reliably differentiate between players?

Independence: Does the metric provide new information?

-->

@@

<!-- [CHECK/EDIT THIS, MAKE SURE WORDING IS CORRECT/NOT CONFUSING] -->

@@

Moreover, we observe statistically significant differences among the pass rush and pass block positions in most cases, while controlling for other variables.
The coefficient estimates for the pass rush position terms reveal that players with positions closer to the line of scrimmage (DTs and NTs) tend to generate less STRAIN (relative to DEs) than those lining up further back, which is consistent with our results in Section \ref{sec:positional}.
For blockers, centers are those that absorb the most pressure on average, more than guards and other offensive positions (compared to the baseline level, tackles).

<!-- [INSERT CONCLUSIONS ABOUT THESE VARIABLES] -->
<!-- [SAY SOMETHING ABOUT POSITION] -->

@@

\begin{table}[ht]
\caption{Intraclass correlation coefficients for the multilevel model for average STRAIN. \label{tab:icc}}
\centering
\begin{tabular}{cccccc}
\hline
\text{ } & Pass rusher & Pass blocker & Defensive team & Offensive team & Residual \\ 
\hline
ICC & 0.0365 & 0.0098 & 0.0134 & 0.0143 & 0.9260 \\ 
\hline
\end{tabular}
\end{table}

@@

In addition to adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the varying intercept distributions for the top ten pass rushers for each position.
We observe similar rankings based on the random intercepts in comparison to the average STRAIN results in Tables \ref{tab:edge} and \ref{tab:interior}, with Myles Garrett (DE), Aaron Donald (DT), Vita Vea (NT), and T.J. Watt (OLB) as the first-ranked pass rusher for their respective position.
We observe different levels of variability across these top players, with only a subset of players with intercept ranges strictly above zero.
This is not necessarily surprising given the limited sample of data.
Additionally, we are ignoring correlated performance between defensive linemen on the same team.
For instance, the top two defensive ends, Myles Garrett and Takkarist McKinley (see the top panel of Figure \ref{fig:fig_rankings_boot}), were teammates on the Cleveland Browns during this period of time.
McKinley is a surprising name on our list and we suspect that his high rank is mostly due to playing on the same defensive unit with Garrett, who is well-known as an elite pass rusher.
We leave the exploration of adjusting for teammates for future work.

[TODO: edit the above? McKinley still there but we did control for team effects]

```{r}
#| label: fig_rankings_boot
#| fig.cap: "Distributions of player effects for top 10 pass rushers in each position (DE, DT, NT, OLB) by median varying intercept."
#| fig.width: 6.3
#| fig.height: 5
source(here("scripts", "10_multilevel_mod.R"))
fig_rankings_boot
```

>>>>

```{r}
#| label: fig_crosby_curves
#| fig.cap: "Changes in STRAIN, velocity, and distance from the quarterback for Maxx Crosby over the course of a successful pass rush play that results in a sack. The STRAIN for other pass rushers throughout this play is also displayed."
#| fig.width: 4.8
#| fig.height: 2.7
source(here("scripts", "02_fig_crosby_curves.R"))
fig_crosby_curves
```

@@

We observe that for the first two seconds, Crosby is being blocked by a Denver offensive lineman and unable to get close to the quarterback, hence the corresponding STRAIN values are virtually zero. 
Suddenly at around three seconds after the snap, the STRAIN for Crosby starts to increase after the Raiders defensive end is freed up and charges towards Bridgewater. 
At 4 seconds after the snap, Crosby's STRAIN is 2.30, which means at his current moving rate, it will take Crosby about 0.43 (1/2.30) seconds to make the distance between him and the quarterback 0 (i.e. essentially sack the quarterback).
This matches well with the final outcome of the play, as the sack takes place at the very last frame (4.4 seconds) where the estimated STRAIN for Crosby reaches its peak at 3.96.

@@

```{r}
#| label: fig_crosby_curves_unsucess
#| fig.cap: "Changes in STRAIN, velocity, and distance from the quarterback for Maxx Crosby over the course of an unsuccessful pass rush play. The STRAIN for other pass rushers throughout this play is also displayed."
#| fig.width: 4.8
#| fig.height: 2.7
source(here("scripts", "09_fig_play_unsuccess.R"))
fig_crosby_curves_unsucess
```

<!-- ```{r} -->
<!-- #| label: fig_features_combined -->
<!-- #| fig.cap: "Changes in STRAIN, velocity, and distance from the quarterback for Maxx Crosby over the course of a successful pass rush play that results in a sack (left) and an unsuccessful pass rush play which results in a long pass attempt (right). The STRAIN for other pass rushers throughout each play is also displayed." -->
<!-- #| fig.width: 6 -->
<!-- #| fig.height: 3.6 -->
<!-- source(here("scripts", "15_fig_features_combined.R")) -->
<!-- fig_features_combined_bw -->
<!-- ``` -->

@@

```{r}
#| label: fig_pos_curves
#| fig.cap: "STRAIN curves for different positions. Edge rushers (outside linebackers and defensive ends) tend to generate more STRAIN than interior rushers (defensive tackles and nose tackles). The dotted gray line represents the average STRAIN curve for all players without accounting for position."
#| fig.width: 4.5
#| fig.height: 3
source(here("scripts", "03_fig_pos_curves.R"))
fig_pos_curves
```

@@

Figure \ref{fig:fig_pos_curves} displays the average STRAIN by position for the first 40 frames (4 seconds) after the snap.
For each position, the curve is based on the average of the observed STRAIN at each frame across all plays and players within the position.
We observe a clear difference in STRAIN between edge rushers (outside linebackers and defensive ends) and interior linemen (defensive tackles and nose tackles).
Specifically, edge rushers have higher STRAIN than interior linemen on average, as they are more easily able to approach the quarterback on the edge of the pocket versus, for instance, a nose tackle attacking the line head on.

@@

Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size, as discussed in Section \ref{sec:defn}.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass rushers (nose tackles and defensive tackles) separately from edge rushers (outside linebackers and defensive ends).
Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season.
The tables also consist of the total number of hits, hurries, and sacks (determined from PFF scouting data) for each defender.

@@

@@

@@

Moreover, we observe statistically significant differences among the pass rush and pass block positions in most cases, while controlling for other variables.
The coefficient estimates for the pass rush position terms reveal that players with positions closer to the line of scrimmage (defensive tackles and nose tackles) tend to generate less STRAIN (relative to defensive ends) than those lining up further back, which is consistent with our results in Section \ref{sec:positional}.
For blockers, centers are those that absorb the most pressure on average, more than guards and other offensive positions (compared to the baseline level, tackles).

@@

\begin{table}
\caption{Intraclass correlation coefficients for the multilevel model for average STRAIN. \label{tab:icc}}
\centering
\begin{tabular}{cccccc}
\hline
\text{ } & Pass rusher & Pass blocker & Defensive team & Offensive team & Residual \\ 
\hline
ICC & 0.0365 & 0.0098 & 0.0134 & 0.0143 & 0.9260 \\ 
\hline
\end{tabular}
\end{table}

@@

In addition to adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the varying intercept distributions for the top ten pass rushers for each position.
We obtain similar rankings based on the random intercepts in comparison to the average STRAIN results in Tables \ref{tab:edge} and \ref{tab:interior}, with Myles Garrett (defensive end), T.J. Watt (outside linebacker), Aaron Donald (defensive tackle), and Vita Vea (nose tackle) as the first-ranked pass rusher for their respective position.
We observe different levels of variability across these top players, with only a subset of players with intercept ranges strictly above zero.
This is not necessarily surprising given the limited sample of data.

[TODO: edit the below? McKinley still there but we did control for team effects]

Additionally, we are ignoring correlated performance between defensive linemen on the same team.
For instance, the top two defensive ends, Myles Garrett and Takkarist McKinley (see the top panel of Figure \ref{fig:fig_rankings_boot}), were teammates on the Cleveland Browns during this period of time.
McKinley is a surprising name on our list and we suspect that his high rank is mostly due to playing on the same defensive unit with Garrett, who is well-known as an elite pass rusher.
We leave the exploration of adjusting for teammates for future work.

```{r}
#| label: fig_rankings_boot
#| fig.cap: "Distributions of player effects for top 10 pass rushers in each position (defensive end, outside linebacker, defensive tackle, nose tackle) by median varying intercept."
#| fig.width: 6.3
#| fig.height: 6
source(here("scripts", "10_multilevel_mod.R"))
fig_rankings_boot
```
